---
title: "What tells you that you have a heart disease?"
subtitle: "An analysis of Heart Disease indicators"
author: 
  - Puyu Liu (1005092971)
thanks: "Code and data are available at: https://github.com/liupuyu2/Puyu-Liu."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "This paper analyzed the data of personal key indicators of heart disease. The result indicates that age, sex, BMI, etc are correlated to having a heart disease.  Based on the analysis, this paper consists of a linear regression analysis, with the use of the programming language R. Keywords: Heart Diseases, Physical health, Mental Health, kaggle, age & sex, stroke, Smoking, Alcohol Drinking"
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
library(knitr)
library(janitor)
library(dplyr)
```

# Introduction

As heart diseases have caused most deaths for people in the US, this dataset include about 20 variables that involve several risk factors of heart disease. Around half of all americans have at least one heart disease indicator: blood pressure, high cholesterol, smoking, alcohol drinking, diabete and obesity (high BMI), physical and mental health. Hopefully we can derive a regression analysis to use the logistic regression model to make future predictions for population with heart disease risk factors. Computational developments have a great effect on detecting the early stage of some typical heart diseases.

The paper starts with an overview of the data with 300 thousands observations. And also, we will look into the correlation of every variable then try to conduct a regression analysis. Before that, the dataset should be properly cleaned and outliers and leverage should be well taken care of. After processing the data, we will be using the ANOVA test to determine our variables in our model.



# Data
In this paper we discuss the key indicators of heart disease from BMI, smoking, and other categorical variables. The data set that we used is provided by the CDC whom surveyed by telephone to gather data from residents across the US and I retrieved the organized data set from the Kaggle website. The dataset that we used surveys 400,000 adults whom are asked a series of questions related to their health status, for example "Have you smoked more than 100 cigarettes in your life?" etc. Many variables were documented by the CDC but only relevant variables related to increasing a risk of heart disease were used and cleaned up within our data set. After cleaning, only roughly 320,000 observations were noted and had relevancy towards directly or indirectly influencing heart disease. 


```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Read the data
dat <- read_csv("/cloud/project/inputs/data/heart_2020_cleaned.csv")
```

```{r}
summary(dat)
```

```{r}
dat=dat %>%
  mutate(y = ifelse(HeartDisease == "No",0,1))
dat=dat %>%
  mutate(smoke = ifelse(Smoking == "No",0,1))
dat=dat %>%
  mutate(drink = ifelse(AlcoholDrinking == "No",0,1))
dat=dat %>%
  mutate(stro = ifelse(Stroke == "No",0,1))
dat=dat %>%
  mutate(dif = ifelse(DiffWalking == "No",0,1))
dat=dat %>%
  mutate(smoke = ifelse(Smoking == "No",0,1))
dat=dat %>%
  mutate(physic = ifelse(PhysicalActivity == "No",0,1))
dat=dat %>%
  mutate(smoke = ifelse(Smoking == "No",0,1))
```



# Model

Y indicates if this person have heart disease.
pi is the probability of having heart disease.
p/1-p is the odds ratio.
X1 - X7 is .....
$$Y_i\sim Bernuolli(p_i)$$

$$log(\frac{p_i}{1-p_i})=\beta_0+X_1\beta_1+X_1\beta_1+X_1\beta_1+X_1\beta_1+X_1\beta_1+X_1\beta_1+X_1\beta_1$$


```{r}
fit = glm(y~BMI+Smoking+AlcoholDrinking+PhysicalHealth+MentalHealth+AgeCategory+SleepTime,family = binomial, data = dat)
```



```{r}
exp(fit$coefficients)
```

```{r}
summary(fit)
```



# Results

From the table above, as age increases, we see that the coefficients for these variables becomes higher and this would mean that the people in a higher age group will be most at risk to heart disease. 

# Discussion

## First discussion point


## Second discussion point

## Third discussion point

## Weaknesses and next steps



\newpage

\appendix

# Appendix {-}


# Additional details


\newpage


# References


